\documentclass[thesis=M,english]{FITthesis}[2012/10/20]

% graphics files inclusion
\usepackage{graphicx}
% subfigures
\usepackage{subfig}
% advanced maths
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
% directory tree visualisation
\usepackage{dirtree}
% tikz package and the mindmap library
\usepackage{tikz}
\usetikzlibrary{arrows,mindmap,shapes.geometric}

\department{Department of Applied Mathematics}
\title{Neural Networks Based Domain Adaptation in Spectroscopic Sky Surveys}
\authorGN{Ondřej}
\authorFN{Podsztavek}
% author's name without academic degrees
\author{Ondřej Podsztavek}
\authorWithDegrees{Bc. Ondřej Podsztavek}
\supervisor{RNDr. Petr Škoda, CSc.}
\acknowledgements{
	I would like to thank Dr. Petr Škoda for his enthusiasm, domain expertise and encouragemenent throughout my research.
	This project would have been impossible without the support of Professor Tvrdík.
	Thanks to my girlfriend Kateřina. You supported me greatly.
	Thanks to Petr and Vojtěch for their friendship in the lab.

	Furthermore, I would like to acknowledge the Sloan Digital Sky Survey and~the Large Sky Area Multi-Object Fiber Spectroscopic Telescope for providing their data.

	Funding for the Sloan Digital Sky Survey IV has been provided by~the~Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS-IV acknowledges
support and resources from the Center for High-Performance Computing at
the University of Utah. The SDSS web site is \url{www.sdss.org}.

Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope) is a National Major Scientific Project built by~the~Chinese Academy of Sciences. Funding for the project has been provided by~the~National Development and Reform Commission. LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences.
}
\abstractEN{
	We present an analysis of the impact of neural-based domain adaptation in~astronomical spectroscopy.
	Domain adaptation addresses the~problem of~applying prior knowledge to a new data of interest.
	Therefore, we selected a~problem~of quasar identification in the Large Sky Area Multi-Object Fiber Spectroscopic Telescope survey using labelled data from the Sloan Digital Sky Survey.
	We choose to experiment with four neural models for domain adaptation:
	Deep Domain Confusion, Deep Correlation Alignment, Domain-Adversarial Network and Deep Reconstruction-Classification Network.
	However, our experiments reveal that these model cannot improve classification performance in comparison to a convolutional neural network
	that does not consider domain adaptation.
	Using dimensionality reduction, statistics of~the~selected methods and misclassifications,
	we show that the domain adaptation methods are not robust enough to be applied to the complex and dirty astronomical data.
}
% TODO czech abstract
\abstractCS{
	Představujeme analýzu dopadu doménové adaptace založené na~neuronových sítích v astronomické spektroskopii.
	Doménové adaptace řeší problém použití dříve získaných znalostí na nová data.
	Analýzu ukazujeme na problému identifikace kvasarů v přehlídce Large Sky Area Multi-Object Fiber Spectroscopic Telescope pomocí anotovaných dat z přehlídky Sloan Digital Sky Survey.
	Pro experimenty jsme vybrali čtyři modely založené na neuronových sítích pro doménovou adaptaci:
	Deep Domain Confusion, Deep Correlation Alignment, Domain-Adversarial Network and Deep Reconstruction-Classification Network.
	Výsledky experimentů ukázaly, že tyto modely nejsou schopné vylepšit klasifikační přesnost v porovnání s konvoluční neuronovou sítí,
	která doménovou adaptaci nebere v potaz.
	S využitím redukce dimensionality, statistik zmíněných metod a chyb v klasifikaci ukazujeme,
	že zvolené metody doménové adaptace nejsou dostatečně robustní, abychom je mohly aplikovat na komplexní a nevyčištěná astronomická data.
}
\placeForDeclarationOfAuthenticity{Prague}
\keywordsEN{
	domain adaptation,
	neural networks,
	deep learning,
	astronomical spectroscopy,
	astronomy
}
\keywordsCS{
	doménová adaptace,
	neuronové sítě,
	hluboké učení,
	astronomická spektroskopie,
	astronomie
}
% select according to the desired license (integer 1-6)
\declarationOfAuthenticityOption{2}
% optional thesis URL
%\website{}

\begin{document}

\chapter{Introduction}

In this thesis, we ask the question of what is the impact of domain adaptation based on neural networks in astronomical spectroscopy. Having two different spectroscopic sky surveys, are we able to extract knowledge from one and~apply it to another with a neural network. To be concrete, we will experiment with the identification of quasars in the Large Sky Area Multi-Object Fiber Spectroscopic Telescope survey with quasars identified in the Sloan Digital Sky Survey.

Previous research in machine learning has shown that domain adaptation can overcome the problem of analysing data from different distributions in one experiment while having a sufficient amount of data. Nowadays, astronomy is facing an avalanche of data, and exponential growth of data transforms science in general. The vast data sources provide immense potential for discoveries. However, we need to adapt existing methods or develop new sophisticated automatic approaches for data analysis. Therefore, we see potential in domain adaptation by exploiting knowledge extracted in previous problems.

To explore the potential of domain adaptation based on neural networks, we set the goal of this thesis to be the analysis of the impact of domain adaptation using data from the Sloan Digital Sky Survey for identification of~quasars in the Large Sky Area Multi-Object Fiber Spectroscopic Telescope survey.

In the second chapter~\ref{data_chapter}, we firstly introduce astronomical spectroscopy, describe the aforementioned spectroscopic surveys, and we demonstrate why they are suitable for our problem.
In chapter~\ref{da_chapter}, we provide a summary of~the~theory of domain adaptation in the context of the machine and transfer learning,  survey domain adaptation based on neural network and give an~overview of applications of domain adaptation in astronomy.
In the chapter~\ref{exp_chapter} with experiments, we analyse the impact of domain adaptation on four models based on neural networks. We suggest future improvements and conclude this thesis in the last chapter~\ref{conclusion}.

\include{spec_sky_surveys}
\include{domain_adaptation}
\include{experiments}

\chapter{Conclusion}
\label{conclusion}

Our goal was to analyse the impact of domain adaptation based on neural networks in spectroscopic sky surveys.
Therefore, we firstly introduced astronomical spectroscopy.
Then, we defined quasars because their identification is the task of our domain adaptation setting.
To have the domain adaptation setting complete,
we introduced the SDSS as the source domain and the~LAMOST survey as the target domain.
We showed that these two domains are suitable for domain adaptation
because of their different instrument and distribution of observations.

Next, we surveyed domain adaptation.
Firstly, we put it into the context of the machine and transfer learning,
and we formally defined it.
Secondly, we distinguished the shallow and deep domain adaptation.
The deep domain adaptation category is based on neural networks,
thus, further, we focused on it, and we presented its three subcategories:
discrepancy-based, adversarial-based and reconstruction-based domain adaptation.
We selected four appropriate methods DDC, Deep CORAL, DANN and DRCN.
DDC and Deep CORAL are representatives of the discrepancy-based category, DANN of adversarial-based methods and DRCN is based on reconstruction.
We explained why we think they are fundamental to experiment with
and give the~intuition and mathematics behind them.

The thing left was to explore their impact of the quasar identification task from astronomical spectra.
Therefore, we briefly introduced data preparation
(actually we spend a significant amount of time with data preparation).
We followed with dimensionality reduction using PCA, t-SNE and UMAP
that demonstrated a visually notable distribution discrepancy between the source and target domain.
To have a baseline model to compare the deep domain adaptation methods to,
we used the LeNet-5 CNN trained only on the source domain.
Its result states that it can be robust classifier on the source domain
but is not able to classify well on the target domain as we expected due to distribution difference.
We continued with the central experimental part of~our thesis.
We applied the four deep domain adaptation methods to~the~data using both the source and target data.

However, the experiments proved that none of DDC, Deep CORAL, DANN and DRCN could achieve a significant improvement on the target data.
That happens, even though dimensionality reduction indicate domain discrepancy
and, for example, the training of DDC keeps MMD of the two distribution low in comparison when the MMD loss is not enforced.
Therefore, we investigated the data in order to analyse
what causes the inability of deep domain adaptation methods to improve performance.
Our astronomical dataset contains a lot of problematic samples
(for example, not identified quasars or bad spectra).
On the other hand, the original deep domain adaptation methods are based on well-prepared and clean data.

That is a big issue for the application of such method to scientific data.
As we have shown, astronomy provides such a volume of data that is impossible to make clean and labelled.
Therefore, we need either robust machine learning algorithms or precise procedures
that clean data possibly also based on machine learning.
However, by cleaning the data, we might lose some interesting objects with~strange physical properties.

\section{Future Plans}

From the research point of view, the result is positive as it opens discovery opportunities because algorithms verified on the well-formed academical dataset like MNIST cannot be directly applied to complex and dirty scientific data. However, we see the future as full of such complex data that we are unable to clean.

Therefore, in future, we plan to make a more detailed statistical analysis of our results and try to make direct consequences for deep domain adaptation methods. The improvements can be either in the form of new robust architectures or novel learning algorithms. Another possibility is to focus on~automatic preprocessing algorithms based on machine learning that will be able to filter the data so that they can work with the original deep domain adaptation methods.

\bibliographystyle{iso690}
\bibliography{references}

\appendix

\include{spectra_appendix}

\chapter{Acronyms}

\begin{description}
	\item[AGN] Active galactic nuclei
	\item[APOGEE] Apache Point Observatory Galactic Evolution Experiment
	\item[BOSS] Baryon Oscillation Spectroscopic Survey
	\item[CDD] Charge-coupled device
	\item[CORAL] Correlation Alignment
	\item[CNN] Convolutional neural network
	\item[DA] Domain adaptation
	\item[DANN] Domain-Adversarial Neural Networks
	\item[DDC] Deep Domain Confusion
	\item[DeCAF] Deep Convolutional Activation Feature
	\item[DRCN] Deep Reconstruction-Classification Network
	\item[EM] Electromagnetic
	\item[FITS] Flexible Image Transport System
	\item[FWHM] Full width at half maximum
	\item[GPU] Graphics processing unit
	\item[GRL] Gradient reversal layer
	\item[IID] Independent and identically distributed 
	\item[LAMOST] Large Sky Area Multi-Object Fiber Spectroscopic Telescope
	\item[QSO] Quasar, Quasi-stellar object
	\item[SDSS] Sloan Digital Sky Survey
	\item[SGD] Stochastic gradient descent
	\item[SNR] Signal-to-noise ratio
	\item[t-SNE] t-Distributed Stochastic Neighbor Embedding
	\item[UMAP] Uniform Manifold Approximation and Projection
\end{description}

\chapter{Contents of Enclosed CD}

\begin{figure}
\dirtree{%
    .1 README.md\DTcomment{file with CD contents description}.
    .1 thesis.pdf\DTcomment{thesis text in PDF format}.
    .1 src\DTcomment{directory of source codes}.
    .2 latex\DTcomment{directory of \LaTeX{} source codes of the thesis}.
    .2 experiments\DTcomment{directory of source codes of experiments}.
}
\end{figure}

\end{document}
